Assesing Health Care Coverage and Access Utilizing the National Health Interview Survey

### 2017-03-08 paper/paper.Rmd
update on final paper and final resource

---
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: '~/Google Drive/INFO498C/INFO498Final/paper/svm-latex-ms.tex'
title: "Assesing Health Care Coverage and Access Utilizing the National Health Interview Survey"
author:
- name: John Diego
  affiliation: University of Washington
- name: Warren Wakuzawa
  affiliation: University of Washington
- name:  Adrian Santiago
  affiliation: University of Washington
abstract: "The National Health Interview Survey (NHIS) is the nation’s largest in-person household health survey. It has been conducted annually since 1957 by the National Center for Health Statistics (NCHS), which is a part of the Centers for Disease Control and Prevention (CDC). A broad range of topics are covered but we will specifically focus on healthcare access and expenditure across a multitude of factors such as employer information, race, and academic background."
keywords: "healthcare, NHIS, ACA, medicaid, medicare"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 10pt
# spacing: double
---

@@

@@

@@

@@

>>>>

---
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: '~/Desktop/INFO498Final/paper/svm-latex-ms.tex'
title: "Assesing Health Care Coverage and Access Utilizing the National Health Interview Survey"
author:
- name: John Diego
  affiliation: University of Washington
- name: Warren Wakuzawa
  affiliation: University of Washington
- name:  Adrian Santiago
  affiliation: University of Washington
abstract: "The National Health Interview Survey (NHIS) is the nation’s largest in-person household health survey. It has been conducted annually since 1957 by the National Center for Health Statistics (NCHS), which is a part of the Centers for Disease Control and Prevention (CDC). A broad range of topics are covered but we will specifically focus on healthcare access and expenditure across a multitude of factors such as employer information, race, and academic background."
keywords: "healthcare, NHIS, ACA, medicaid, medicare"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 10pt
# spacing: double
---

@@

Because the dataset was so large, it was difficult to manage and processing the data was time-consuming.  In order to fully combat this, we used the dplyr package in order to modify the dataset we would be using in such a way that only incorporated the columns we needed to address the question at hand.  After, we then filtered the data so that we would have only the information we needed to answer the questions, while weeding out “empty” data such as “not ascertained,” “didn’t know,” as well as “refused.”  When we were certain of the dataset we needed that only contained relevant information to our questions, we then utilized the ggplot2 R package in order to create well-labeled graphs that visually summarized our results.  From the graphs, we were able to draw reasonable conclusions and generate an analysis of why the data was shaped in the particular way it was.

@@

In respects to the National Health Interview Survey, we wanted to research what percentage of the population had healthcare coverage. We used the “NOTCOV” column of the survey to retrieve the amount of individuals that had coverage and those who did not have healthcare coverage. We do want to mention that there was a total of 1102 individual that responded with “Don’t know”. WE made the decision to exclude these individuals from our data since they do not represent your targeted population.

When comparing just the proportions of those who had healthcare coverage to those who lacked coverage, there was a clear transformation. Only 10.23% of the individuals that participated in this survey did not have healthcare coverage. While 89.76% of these individuals did have coverage.

@@

```{r hi2015, echo=FALSE}
barplot(h, names.arg = m, xlab = "Healthcare Coverage for 2015", 
        ylab="Percentage (%)", col='blue', main = "Percentage of the Respondents that Has Healthcare Coverage")
```

**Figure 5. Percentage of the Respondents that Has Healthcare Coverage**

```{r hi2015employer, echo=FALSE}
barplot(h2, names.arg = m2, xlab = "Healthcare Coverage for 2015", 
        ylab="Percentage (%)", col='blue', main = "Percentage of the Respondents that Has Healthcare Coverage")
```
**Figure 6. Percentage of the Respondents that Has Healthcare Coverage**

@@

## **Health Care Access**
A huge discussions these recent weeks has been Obama care. Some view it was a total disaster while others suggest that it works but it is not perfect. Our president has spoken that this plan is terrible and this inspired us to consider the what a reality would be like if there was no more Obama care. As a group we believed that without government help, individuals would have to resort to other sorts of health care plans. We decided to focus on employers to study if there is any significant correlation between employers and health care coverage. We do have to note that this information lacked the participation from 42728 individuals. This number was subtracted from the group that did responded as we were interested in the population that did have some sort of coverage.

As our prior graph indicates, about 89.76% of the surveyed population expressed some sort of healthcare coverage. Out of this population we calculated that 78.69% of population that has some sort of healthcare coverage attains it through an employer. This is only about a 11.07% difference suggesting that the way more people get their healthcare coverage through their workforce. However we did not count those individuals who obtained coverage through their own self-employed business, a union, or even those who were uncertain about their response.

After analyzing these findings, it is our suggestion that there is some relationship between employers and a person's’ health care coverage. But must point out that it is not very clear to understand how strong this correlation is or to even identify in what specific ways these aspects have an influence on each other. The 11% difference does portray the idea that everyone is obtaining healthcare coverage from their employers. But it is crucial to realize that the portion that missed this question represents, in quantity,  consists of more than half the population that did decide to answer. This great lack of representation from the population is too big to consider a strong and clear correlation between employers and health care coverage.
